---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wateRtemp

<!-- badges: start -->
<!-- badges: end -->

A machine learning toolbox for river water temperature prediction.

## Installation

You can install the released version of wateRtemp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("MoritzFeigl/wateRtemp")
```
## Example

This is a basic example of how to preprocess data and how to apply different machine learning algorithms using wateRtemp 
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(wateRtemp)
```

At first it is necessary to create a folder with the name of the catchment or study area in which the preprocessed data and results will be stored. For that we create here a dummy folder called "Catchment".
```{r example}
dir.create("Catchment")
```

For using watRtemp preprocessing and modelling functions, the necessary data should be available as a data frame. An example data frame is included in wateRtemp and can be called by `data("wt_data")`.
```{r}
# Provide the catchment data as a data frame
data("wt_data")
# it should look like this:
str(wt_data)
```

After loading the necessecery data, we can use the watRtemp preprocessing function to apply feature engineering and data splits. The preprocessed data will be saved in the Catchment folder automatically. 
```{r}
# Preprocess the data
wt_preprocess("Catchment", data = wt_data,
              year_range = c(2007, 2015))
```

Now we are ready to apply our machine learning models. For this example we run a stepwise linear model using the function `wt_lm` with the specification `type = "step"`. 
```{r model_run, warning=FALSE}
wt_lm(catchment = "Catchment",
      data_inputs = "simple", 
      type = "step", 
      user_name = "Example")
```
Similar to linear models, we can easily apply other type of machine learning models on the preprocessed data (including hyperparemeter optimization):

* `wt_randomforest` for random forests,

* `wt_xgboost` for XGBoost (Extreme Gradient Boosting),

* `wt_ann` for feed forward neural networks,

* `wt_rnn` for recurrent neural networks: LSTMs and GRUs

```{r echo=FALSE, message=FALSE, warning=FALSE}
unlink("Catchment", recursive = T)
```
