---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wateRtemp

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/186992552.svg)](https://zenodo.org/badge/latestdoi/186992552)
<!-- badges: end -->

A machine learning toolbox for daily mean river water temperature prediction.

# Overview
watRtemp includes 6 machine learning models with already implemented bayesian hyperparameter optimization. The main functions are:

* `wt_preprocessing()` for preprocessing data for the machine learning models,

* `wt_lm()` for muliple regression and step-wise linear regression,

* `wt_randomforest()` for random forests,

* `wt_xgboost()` for XGBoost (Extreme Gradient Boosting),

* `wt_fnn()` for feedforward neural networks,

* `wt_rnn()` for recurrent neural networks: LSTMs and GRUs

Additionally, a prepared synthetic data set for testing wateRtemp functionalities is included and can be used by running:
```{r}
data("test_catchment")
```

## Installation

You can install the released version of wateRtemp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("MoritzFeigl/wateRtemp")
```
## Example

This is a basic example of how to preprocess data and how to apply different machine learning algorithms using wateRtemp.
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(wateRtemp)
```

For using watRtemp preprocessing and modelling functions, the necessary data should be available as a data frame. Example data frames are included in wateRtemp and can be called by `data("test_catchment")`.
```{r}
# Provide the catchment data as a data frame
data("test_catchment")
# it should look like this:
summary(test_catchment)
```

After loading the necessary data, we can use the watRtemp preprocessing function to apply feature engineering and data splits. The preprocessed data will be saved in the Catchment folder automatically. 
```{r}
# Preprocess the data
wt_preprocess(test_catchment, nlags = 4, training_test_fractions = c(0.8, 0.2))
```

After preprocessing, the corresponding training and test datasets are stored in the catchment folder and can be loaded for using them in the models. 
```{r}
# Preprocess the data
train_data <- feather::read_feather("test_catchment/train_data.feather")
test_data <- feather::read_feather("test_catchment/test_data.feather")
```

Now we are ready to apply our machine learning models. For this example we run the most simple model available in wateRtemp: a multiple regression model using the function wt_lm(). 
```{r model_run, warning=FALSE}
wt_lm(train_data, test_data, "test_catchment", "LM", "repCV", "standard_LM")
```
Similar to multiple regression, we can easily apply other type of machine learning models on the preprocessed data. All models in include Bayesian hyperparemeter optimization and automatically storing of results and trained models.

* `wt_lm(type = "stepLM")` for step-wise linear regression,

* `wt_randomforest()` for random forests,

* `wt_xgboost()` for XGBoost (Extreme Gradient Boosting),

* `wt_ann()` for feed forward neural networks,

* `wt_rnn()` for recurrent neural networks: LSTMs and GRUs

